\LARGE \bf Machine Translation for Tibetan Buddhist Texts

### 2021-04-16 paper/main.tex
cited image
Update main.tex
Fixed typos in abstract

This paper’s presents our attempt at training a model focusing on the translation of religious Tibetan texts to the English language. Tibetan is an extremely low-resource language and presents its own unique challenges in every stage toward the development of a high accuracy model. We look to explore ethical considerations of translating religious texts, provide an in-depth view of our procedure and follow up with discussions, results and conclusions. In support, we utilize various resources including conversations with a native speaker and the exploration of open source parallel Tibetan corpuses.

@@

\subsection{Model Design}
We used a 6 layer Transformer model that used a multi headed attention mechanism and a switchout transform. Each layer of the transformer model can be divided into an encoder and decoder layer. The encoding layer consists first of a self attention layer, an attention aspect which focuses on the encoded sentence, and then outputs to a feed forward layer. The multi part of the multi headed attention means that each head will encode the information in parallel and then the results will be concatenated and passed to the decoder. Figure one demonstrates this process.
\includegraphics[scale=.4]{Transformer.png}
\caption{\\Figure 1: Overview of a Transformer model}
\\
\par
To prevent overfitting and to augment the data, we added in a switchout transform which will randomly switch out words in the source and target sentence. This also has the added advantage of augmenting the data even further which is especially important for our problem as religious Tibetan has a relatively small corpus, around 30 thousand words. We also attempted to use masking, however we found it to be ineffective. Masking hides tokens from the model by replacing the tokens with a <mask> token. Tibetan tokens are not separated in the same way that latin languages separate their tokens. For example,
a line in our test corpus translated to “Having the altruistic intention of excellent bodhicitta.” However, the Tibetan word is only one token. Therefore, the mask would not help the model learn at all, as the entire sentence would be marked as <mask> and there would not be enough context for our model to discover the meaning. As a result, we did not use any form of masking in our model. Switchout does not create the same problem as their is always a token to take its place. \par
After implementing our transformer model with switchout we tuned hyper parameters to maximize our Transformer model. We manually tuned the following hyperparameters: dropout, label smoothing, rnn size, the number of transformer layers and the number of heads of the attention model. The final configuration used can be found in the appendix. \par
We used four different metrics to evaluate our results: accuracy, predicted score, perplexity and bilingual evaluation understudy (BLEU). The validation accuracy tells us how well we are learning our training data, the predicted score tells us the average log likelihood for each word, the perplexity tells us how well the model predicts the data and the bleu score evaluates the quality of the translation. Together, these metrics provide us with a decent picture of how well our model is performed.

@@

>>>>

This paper presents our attempt at training a model focusing on the translation of religious Tibetan texts to the English language. Tibetan is an extremely low-resource language and presents its own unique challenges in all stages of the development of a high accuracy machine translation model. In this paper, we explore the ethical considerations of translating religious texts, provide an in-depth view of our model creation procedure and follow up with discussions, results and conclusions. In support, we utilize various resources including scheduling conversations with a native speaker and the exploration of open source parallel Tibetan corpora.

@@

\subsection{Model Design}
We used a 6 layer Transformer model that used a multi headed attention mechanism and a switchout transform. Each layer of the transformer model can be divided into an encoder and decoder layer. The encoding layer consists first of a self attention layer, an attention aspect which focuses on the encoded sentence, and then outputs to a feed forward layer. The multi part of the multi headed attention means that each head will encode the information in parallel and then the results will be concatenated and passed to the decoder. Figure one demonstrates this process.
\includegraphics[scale=.4]{Transformer.png}
\caption{\\Figure 1: Overview of a Transformer model \cite{11}}
\\
\par
To prevent overfitting and to augment the data, we added in a switchout transform which will randomly switch out words in the source and target sentence. This also has the added advantage of augmenting the data even further which is especially important for our problem as religious Tibetan has a relatively small corpus, around 30 thousand words. We also attempted to use masking, however we found it to be ineffective. Masking hides tokens from the model by replacing the tokens with a <mask> token. Tibetan tokens are not separated in the same way that latin languages separate their tokens. For example,
a line in our test corpus translated to “Having the altruistic intention of excellent bodhicitta.” However, the Tibetan word is only one token. Therefore, the mask would not help the model learn at all, as the entire sentence would be marked as <mask> and there would not be enough context for our model to discover the meaning. As a result, we did not use any form of masking in our model. Switchout does not create the same problem as their is always a token to take its place. \par
After implementing our transformer model with switchout we tuned hyper parameters to maximize our Transformer model. We manually tuned the following hyperparameters: dropout, label smoothing, rnn size, the number of transformer layers and the number of heads of the attention model. The final configuration used can be found in the appendix. \par
We used four different metrics to evaluate our results: accuracy, predicted score, perplexity and bilingual evaluation understudy (BLEU). The validation accuracy tells us how well we are learning our training data, the predicted score tells us the average log likelihood for each word, the perplexity tells us how well the model predicts the data and the bleu score evaluates the quality of the translation. Together, these metrics provide us with a decent picture of how well our model is performed.

@@

\bibitem{c11} Doshi, Ketan (2021). Transformers Explained Visually (Part 3): Multi-head Attention, deep dive. https://towardsdatascience.com/transformers-explained-visually-part-3-multi-head-attention-deep-dive-1c1ff1024853
